Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-02-10 trunk/constantRate/ConstCompFuzzExt.tex
cleaning up claim d.3
minor edits to intro

Our first construction requires high individual entropy of a constant fraction of symbols, but permits symbols to be dependent.  Our second construction requires  a constant fraction of symbols to have a constant amount of entropy conditioned on prior symbols. The constructions can be implemented efficiently based on number-theoretic assumptions or assumptions on cryptographic hash functions.
\bnote{more about what we do}

@@

\paragraph{Limitations of Entropy-Based Approaches}
\bnote{shorten this heading?}
Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess some $w'$ within distance $t$ of $w$ they will be able to easily distinguish $r$ from uniform by running $\rep$. If $t$ is larger, then $\rep$ tolerates more $w'$ values, so the adversary's job is easier. In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then the \emph{entire distribution} of $w$ may have min-entropy $m$ and be contained in a single ball.  Then the adversary can always produce the key by running $\rep$ with the center of the ball as $w'$.\bnote{was told this wasn't clear enough}

@@

\medskip
The remainder of this paper is organized as follows: we cover notation and background on obfuscation and error correcting codes in \secref{sec:preliminaries}, describe computational fuzzy extractors in \secref{sec:fuzzy extractors}, and present our two constructions in Sections \ref{sec:construction} and \ref{sec:sampling} respectively.

@@

\begin{definition}~\cite[Definition 2.5]{fuller2013computational}
\label{def:comp fuzzy extractor}
Let $\mathcal{W}$ be a family of probability distributions over $\mathcal{M}$. A pair of randomized procedures ``generate'' $(\gen)$ and ``reproduce'' $(\rep)$ is a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-\emph{computational fuzzy extractor} that is $(\epsilon, s_{sec})$-hard with error $\delta$ if \gen and \rep satisfy the following properties:
\begin{itemize}
\item The generate procedure \gen on input $w\in \mathcal{M}$ outputs an extracted string $R\in\{0,1\}^\kappa$ and a helper string $P\in\{0,1\}^*$.
\item The reproduction procedure \rep takes an element $w'\in\mathcal{M}$ and a bit string $P\in\{0,1\}^*$ as inputs.  The \emph{correctness} property guarantees that if $\dis(w, w')\leq t$ and $(R, P)\leftarrow \gen(w)$, then $\Pr[\rep( w', P) = R] \geq 1-\delta$ where the probability is over the randomness of $(\gen, \rep)$.
If $\dis(w, w') > t$, then no guarantee is provided about the output of \rep.
\item The \emph{security} property guarantees that for any distribution $W\in \mathcal{W}$, the string $R$ is pseudorandom conditioned on $P$, that is $\delta^{\mathcal{D}_{s_{sec}}}((R, P), (U_\kappa, P))\leq \epsilon$.
\end{itemize}
\end{definition}
In the above definition, the errors are chosen before $P$: if the error pattern between $w$ and $w'$ depends on the output of $\gen$, then there is no guarantee about the probability of correctness. In both our constructions it is crucial that $w'$ is chosen independently of the outcome of \gen.
\bnote{cut the rest of this paragraph?} The definition of computational fuzzy extractors specifies a family of sources for which the fuzzy extractor works rather than the family of all sources of a given min-entropy $m$.  As discussed in the introduction, when $\Huse\le 0$ some restriction on the source is necessary.  Instead of restricting the class of source distributions, the definition could restrict the correctable errors to errors that are ``likely'' to occur in the source.  We leave this as an open problem.

@@

\begin{corollary}
Let all the variables be as in \thref{thm:sampling}.
When
\[
\dis(w, w')\le \frac{-\log(1-t'/(2\ell))}{2}\frac{\gamma-\eta}{\eta} = \Theta(\gamma/\eta)
\]
and $(c, p)\leftarrow \gen(w)$, then
\[
\Pr[\rep(w', p) = c] \geq 1-\frac{\ell}{|\mathcal{Z}|^\eta} - O(2^{-\ell}) -\delta_{code}= 1-\ngl(n).
\]
That is, \consref{cons:sampling} is correct with overwhelming probability.
\end{corollary}

@@

It suffices to show that there exists an $S$ making $q$ queries such that for some $j\in J, \Hav(W_j | S^{I_{W}(\cdot, \cdot)})\le \log |\mathcal{Z}| - \log q$.  Denote by $w_{j,1}, ..., w_{j,q}$ the $q$ most likely outcomes of $W_j$~(breaking ties arbitrarily).  Then $\sum_{i=1}^q \Pr[W_j = w_{j,i}]\geq q/|\mathcal{Z}|$.  Suppose not, this means that there is some $w_{j,i}$ with probability $\Pr[W_j = w_{j,i}] < 1/|\mathcal{Z}|$.  Since there are $\mathcal{Z} - q $ remaining possible values of $W_j$ for their total probability to be at least $1-q/|\mathcal{Z}|$ at least of these values has probability at least $1/\mathcal{Z}$.  This contradicts the statement $w_{j,1},..., w_{j,q}$ are the most likely values.  Consider $S$ that queries its oracle on $(j, w_{j,1}),.., (j, w_{j,q})$.  Denote by $Bad$ the random variable when $W_j\in \{w_{j,1},.., w_{j,q}\}$  After these queries the remaining min-entropy is at most:

>>>>

Our first construction requires high individual entropy of a constant fraction of symbols, but permits symbols to be dependent.  Our second construction requires  a constant fraction of symbols to have a constant amount of entropy conditioned on prior symbols. The constructions can be implemented efficiently based on number-theoretic assumptions or assumptions on cryptographic hash functions.

@@

\paragraph{Limitations of Known Approaches}

Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess any $w'$ within distance $t$ of $w$, it will be able to easily obtain the true $r$ by running $\rep$.
%If $t$ is larger, then a single guess for $w'$ may be within distance $t$ of more $w$ values, thus increasing adversarial probability of winning. 
In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then there exists a distribution of $w$ of min-entropy $m$  \emph{contained entirely in a single ball}.  For these distributions, an adversary can run $\rep$ on the center of this ball and always learn the key $r$.

@@

\lnote{restricting errors instead of entropy is another way to get around impossibility}

\medskip
\lnote{skip if no space}
The remainder of this paper is organized as follows: we cover notation and background on obfuscation and error correcting codes in \secref{sec:preliminaries}, describe computational fuzzy extractors in \secref{sec:fuzzy extractors}, and present our two constructions in Sections \ref{sec:construction} and \ref{sec:sampling} respectively.

@@

\begin{definition}~\cite[Definition 2.5]{fuller2013computational}
\label{def:comp fuzzy extractor}
Let $\mathcal{W}$ be a family of probability distributions over $\mathcal{M}$. A pair of randomized procedures ``generate'' $(\gen)$ and ``reproduce'' $(\rep)$ is a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-\emph{computational fuzzy extractor} that is $(\epsilon, s_{sec})$-hard with error $\delta$ if \gen and \rep satisfy the following properties:
\begin{itemize}
\item The generate procedure \gen on input $w\in \mathcal{M}$ outputs an extracted string $R\in\{0,1\}^\kappa$ and a helper string $P\in\{0,1\}^*$.
\item The reproduction procedure \rep takes an element $w'\in\mathcal{M}$ and a bit string $P\in\{0,1\}^*$ as inputs.  The \emph{correctness} property guarantees that if $\dis(w, w')\leq t$ and $(R, P)\leftarrow \gen(w)$, then $\Pr[\rep( w', P) = R] \geq 1-\delta$ where the probability is over the randomness of $(\gen, \rep)$.
If $\dis(w, w') > t$, then no guarantee is provided about the output of \rep.
\item The \emph{security} property guarantees that for any distribution $W\in \mathcal{W}$, the string $R$ is pseudorandom conditioned on $P$, that is $\delta^{\mathcal{D}_{s_{sec}}}((R, P), (U_\kappa, P))\leq \epsilon$.
\end{itemize}
\end{definition}
In the above definition, the errors are chosen before $P$: if the error pattern between $w$ and $w'$ depends on the output of $\gen$, then there is no guarantee about the probability of correctness. In both our constructions it is crucial that $w'$ is chosen independently of the outcome of \gen.

@@

%\begin{corollary}
%Let all the variables be as in \thref{thm:sampling}.
%When
%\[
%\dis(w, w')\le \frac{-\log(1-t'/(2\ell))}{2}\frac{\gamma-\eta}{\eta} = \Theta(\gamma/\eta)
%\]
%and $(c, p)\leftarrow \gen(w)$, then
%\[
%\Pr[\rep(w', p) = c] \geq 1-\frac{\ell}{|\mathcal{Z}|^\eta} - O(2^{-\ell}) -\delta_{code}= 1-\ngl(n).
%\]
%That is, \consref{cons:sampling} is correct with overwhelming probability.
%\end{corollary}

@@

It suffices to show that there exists an $S$ making $q$ queries such that for some $j\in J, \Hav(W_j | S^{I_{W}(\cdot, \cdot)})\le \log |\mathcal{Z}| - \log q$.  Let $j\in J$ be some arbitrary element of $J$ and denote by $w_{j,1}, ..., w_{j,q}$ the $q$ most likely outcomes of $W_j$~(breaking ties arbitrarily).  Then $\sum_{i=1}^q \Pr[W_j = w_{j,i}]\geq q/|\mathcal{Z}|$.  Suppose not, this means that there is some $w_{j,i}$ with probability $\Pr[W_j = w_{j,i}] < 1/|\mathcal{Z}|$.  Since there are $\mathcal{Z} - q $ remaining possible values of $W_j$ for their total probability to be at least $1-q/|\mathcal{Z}|$ at least of these values has probability at least $1/\mathcal{Z}$.  This contradicts the statement $w_{j,1},..., w_{j,q}$ are the most likely values.  Consider $S$ that queries its oracle on $(j, w_{j,1}),.., (j, w_{j,q})$.  Denote by $Bad$ the random variable when $W_j\in \{w_{j,1},.., w_{j,q}\}$  After these queries the remaining min-entropy is at most:
